---
title: "caret"
author: "Mark Hagemann"
date: "March 19, 2015"
output: html_document
---

I'm using this project as an excuse to learn the [**caret**](http://topepo.github.io/caret/index.html) package. As I write this, I have only heard of caret, and never so much as installed it on my computer. Here goes!

## 0. Load the package

```{r}
if(!require(caret)) {install.packages("caret"); library(caret)}
vignette("caret")
```

I see that **caret** is something like a front-end for a whole bunch of other packages, which it merely *suggests*. 

```{r}
# install suggested packages:
install.packages("caret", dependencies = c("Depends", "Suggests"))
library(caret)
```

## 0.5: Load the training data:

```{r}
tdata = read.csv("data/training.csv")
```


## 1. Partition data

- Use `createDataPartition()` to make a simple partition.
- `createResample()` makes bootstap resamples.
- `createFolds()` and `createMultiFolds()` make crossvalidation folds
- `createTimeSlices()` for time-series data

I'll start off simply, by making a training/testing partition.

```{r}
part1 = createDataPartition(tdata[[1]], p = 0.8)
str(part1)
```

Hmm. This only returns a vector of indices sampled from `1:nrow(tdata)`. It's not clear how this is useful enough to warrant its own funciton. Wouldn't it be more useful as a logical vector so that I could split the data into train and test by something like:

```{r}
part1_logi = 1:nrow(tdata) %in% part1[[1]]
train1 = tdata[part1_logi, ]
test1 = tdata[!part1_logi, ]
```

More info on this function [here](http://topepo.github.io/caret/splitting.html)

## 2. function: `train`

Read more [here](http://caret.r-forge.r-project.org/training.html)

```{r, eval = FALSE}
?train
```

`train` takes several arguments, one of which is `method`. This can be either a built-in regression or classification method (one of [these options](http://topepo.github.io/caret/bytag.html)), or a custom-built method. The list of built-in methods is long, and appears to include some pretty sophisticated ML methods.

As a first try on this data, I'll use ordered logistic regression via the `polr` method in  `train()`. 

```{r}
tx = train1[2:5]
ty = as.factor(train1[[6]])
try1 = train(x = tx, y = ty, method = "polr")
```

Note that since we're doing classification, `train()` expects `y` to be a factor. 

Also I'm getting an error with `train`. Not very clear, but `warnings()` suggests that it wants more levels in `y`. Maybe *polr* doesn't work as a method in this case? 

Try instead penalized logistic regression, using `method = "plr"`. 

```{r}
try2 = train(x = tx, y = ty, method = "plr")
```

Interesting. This asked if I wanted to install a package ( **stepPlr**). `warnings()` indicate a *computationally singular system*. But it worked. 

```{r}
str(try2, 1)
summary(try2)
```


